Demography and selection since maize domestication

### 2015-05-18 manuscript.tex
minor whinings

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}

@@

\title{Demography and selection since maize domestication}
\author{Timothy M. Beissinger\affil{1}{University of
    California,Davis}, Li Wang\affil{2}{Iowa State University}, Arun
  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center, UC
    Davis}\affil{4}{Center for population biology, UC Davis} }

@@

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from tripsicum, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure \ref{hardSweeps}). Additionally, diversity around maize substitutions
not seen in teosinte was investigated. These sites have the
potential to correspond to recent maize sweeps, occuring after the
split from teosinte. Again, no difference was observed
between diversity around synonymous and nonsynonymous
substitutions \textcolor{red}{(Supplemental figure here)}. Together, these observations suggest that hard sweeps
are not a primary form of selection for either maize or teosinte.

\begin{figure}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_unNeutralized}
\caption{Pairwise diversity surrounding synonymous and nonsynonymous
  substitutions in maize. Observe that the reduction of diversity
  surrounding nonsynonymous subsitutions is no more severe than the
  reduction of diversity surrounding synonymous substitutions. This
  equivalency suggests that hard sweeps are not a primary driver of
  maize diversity.}
\label{hardSweeps}
\end{figure}

@@

\begin{figure*}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_manuscript.png}
\caption{Relative level of diversity versus distance to the nearest
  gene, in maize and teosinte. Two measures of diversity were
  investigated. \textbf{A} displays pairwise
  diversity, which is most influenced by intermediate frequency
  alleles and therefore depicts more ancient evolutionary patterns,
  and \textbf{B} depicts singleton diversity, influenced by rare
  alleles and thus depicting evolutionary patterns in the recent past.}
\label{purify}
\end{figure*}

@@

\begin{figure}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.5\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
\caption{Parameters of domestication as estimated by dadi. Notably,
  the maize effective population size ($N_e$) during the domestication
  bottleneck appears to have consisted of approximately 5.26\%
  the ancestral $N_e$, before recovering two at least 2.98
  times as large as the ancestral $N_e$.}
\label{bottleneck}
\end{figure}

@@

\subsection{Evaluating diversity around substitutions}
To investigate diversity around substitutions, maize and teosinte pairwise diversity was first
calculated in 1,000 kb non-overlapping windows using ANGSD
\cite{korneliussen2014}. This was performed separately for both maize and
teosinte, using the same filters as employed for estimating the
SFS. Next, SNPs and genotypes among maize, teosinte, and tripsicum were called. Tripsicum bam files
were downloaded from (TRIPSICUM FILES), and then all SNPs with a
p-value less than 1e-6 were called using ANGSD. Quality filters were
as the same as before, and genotypes were only called when the
posterior probability was above 0.95. From the set of called SNPs and
genotypes, substitutions between maize and tripsicum, as well as
between teosinte and tripsicum were identified using R \cite{R2014} as all positions with
no more than 20\% missing data for which every maize or teosinte
allele differed from the observed tripsicum allele. At each class of
substitution, effects were estimated using the ensembl variant effects
predictor \cite{mclaren2010}.

@@

\begin{acknowledgments}
Various thankyous will be in order.
\end{acknowledgments}

>>>>

\documentclass{pnastwo}
\usepackage{cite}
\usepackage{color}
\usepackage{pgf}
\usepackage{hyperref}
\newcommand{\jri}[1]{\textcolor{blue}{\emph{#1}} }

@@

\title{Demography and selection since maize domestication}
\author{Timothy M. Beissinger\affil{1}{Dept. of Plant Sciences, University of
    California, Davis, CA, USA}, Li Wang\affil{2}{Iowa State University, Ames, IA, USA}, Arun
  Durvasula\affil{1}{}, Kate Crosby\affil{1}{}, Matthew Hufford\affil{2}{}, \and Jeffrey
  Ross-Ibrarra\affil{1}{}\affil{3}{Genome Center and Center for population biology, University of
    California, Davis, CA, USA} }

@@

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from \emph{Tripsacum}, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure \ref{hardSweeps}). Additionally, diversity around maize substitutions
not seen in teosinte was investigated. These sites have the
potential to correspond to recent maize sweeps, occuring after the
split from teosinte. Again, no difference was observed
between diversity around synonymous and nonsynonymous
substitutions \textcolor{red}{(Supplemental figure here)}. Together, these observations suggest that hard sweeps
are not a primary form of selection for either maize or teosinte.

\begin{figure}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.5\textwidth]{FigsAndFiles/plotDiversity_TvM_Folded2_unNeutralized}
\caption{Pairwise diversity surrounding synonymous and nonsynonymous
  substitutions in maize. \jri{needs larger legend/axis labels. x axis needs to say cM}}
\label{hardSweeps}
\end{figure}

@@

\begin{figure*}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Folded2_manuscript.png} \includegraphics[width=.45\textwidth]{FigsAndFiles/distanceToGene_WithSignificance_Singletons_manuscript.png}
\caption{Relative level of diversity versus distance to the nearest
  gene, in maize and teosinte. Two measures of diversity were
  investigated. \textbf{A} displays pairwise
  diversity, which is most influenced by intermediate frequency
  alleles and therefore depicts more ancient evolutionary patterns,
  and \textbf{B} depicts singleton diversity, influenced by rare
  alleles and thus depicting evolutionary patterns in the recent past. \jri{are these done with latest values we used for curve fitting? should also add the singleton vs pi comparison within species as a supplemental figure}}
\label{purify}
\end{figure*}

@@

\begin{figure}[b]
%\vspace*{.05in}
\centering
\includegraphics[width=.4\textwidth]{FigsAndFiles/DomesticationModel/domesticationModel.pdf}
\caption{Parameters of domestication as estimated by dadi. Notably,
  the maize effective population size ($N_e$) during the domestication
  bottleneck appears to have consisted of approximately 5.26\%
  the ancestral $N_e$, before recovering two at least 2.98
  times as large as the ancestral $N_e$.}
\label{bottleneck}
\end{figure}

@@

\subsection{Evaluating diversity around substitutions}
To investigate diversity around substitutions, maize and teosinte pairwise diversity was first
calculated in 1,000 kb non-overlapping windows using ANGSD
\cite{korneliussen2014}. This was performed separately for both maize and
teosinte, using the same filters as employed for estimating the
SFS. Next, SNPs and genotypes among maize, teosinte, and \emph{Tripsacum} were called. \emph{Tripsacum}\emph{Tripsacum} bam files
were downloaded from \url{}(TRIPSICUM FILES), and then all SNPs with a
p-value less than 1e-6 were called using ANGSD. Quality filters were
as the same as before, and genotypes were only called when the
posterior probability was above 0.95. From the set of called SNPs and
genotypes, substitutions between maize and tripsicum, as well as
between teosinte and \emph{Tripsacum} were identified using R \cite{R2014} as all positions with
no more than 20\% missing data for which every maize or teosinte
allele differed from the observed \emph{Tripsacum} allele. At each class of
substitution, effects were estimated using the ensembl variant effects
predictor \cite{mclaren2010}.

@@

\begin{acknowledgments}
Various thankyous will be in order. \jri{Coop and his lab. funding from NSF Biology of Rare alleles -- need grant number.}
\end{acknowledgments}
